%!TEX root = /home/renaud/Documents/EPL/tfe/latex/tfe.tex
\chapter{Clustering} \label{chap:clustering}
% \section{The stability criterion for graph communities} \label{sec:stability}
The partition of a graph into communities (or clusters) has been widely studied those last two decades. Clustering comes indeed pretty handy to gain insight into the underlying structure of a system represented by a network. In some cases one can even build a simplified functional description of the system based on the clusters. Many partitioning methods have been proposed, each relying on a particular measure to quantify the quality of a community structure. Such methods include normalized cut, ($\alpha$,$\epsilon$) clustering or modularity and its variants and extensions. The reader may refer to \cite{fortunato2010community} for a 2010 survey of the different clustering methods. In this work, we choose the stability approach, which is based on the statistical properties of a dynamical process taking place on the network. This approach was initially presented in \cite{delvenne2010stability} and further expended in \cite{lambiotte2009laplacian} and \cite{delvenne2013stability}. 

The stability method presents a number of advantages. First, it does not require the number of communities to be specified beforehand, ensuring a natural partitioning of the graph. Second, it is flexible in the sense that it does not seek a \textit{unique} optimal partition. Instead, it reveals several community structures, each appearing to be the most relevant at particular values of the Markov time: at a given time scale, natural clusters corresponds to sets of states from which escape is unlikely within that time scale. The stability method provides thus a dynamical interpretation of the partitioning problem. The Markov time acts as an intrinsic resolution parameter, as will be developed shortly. Finally, it is probably the most unifying approach since many of the standard partitioning measures find an interpretation through the stability framework.

In order to compute stability partitions in the next of this work, we make use of Michael Schaub's free software \textit{PartitionStability}. This C++ implementation of the stability method with a \matlab interface is available at \url{https://github.com/michaelschaub/PartitionStability}. It relies on the Louvain algorithm \cite{blondel2008fast} to optimize the stability quality function. This heuristic algorithm has been initially developed for modularity optimization. However one can show that stability can be written as the \textit{modularity} of a time-dependent network evolving under the Markov process \cite{lambiotte2009laplacian}. Hence, the Louvain method can almost straightforwardly be applied to stability optimization.

This chapter is devoted to the explanation of the stability measure, and how to find good clusterings using stability analysis. It acts as a theoretical part intended to cover everything that is needed to make a proper, informed use of the stability toolbox. Notice that the stability measure has initially been presented for discrete times in \cite{delvenne2010stability}. We follow the same approach here: discrete-time stability is developed in the first section of this chapter; it is then extended to continuous time in a second section; finally, a few tools to analyze the robustness of a partition are presented in the third section of the chapter.
\input{inputs/clustering/discretetime}
\input{inputs/clustering/continuoustime}
\input{inputs/clustering/robustness}